Flat-histogram method comparison on 2D Ising Model

### 2020-04-24 ising/ising.tex
went over paper with a fine comb and edited.

\begin{abstract}
  We examine the convergence of the Stochastic Approximation with a Dynamic
  update factor (SAD) algorithm applied to the 2D Ising model. Comparison with
  SAMC and WL methods show that SAD performs robustly and without user input
  knowledge of an energy range. We also compare SAD with a final production run WL simulation with fixed weights which is the first of its kind recorded. Since SAD is more powerful in the common case in which the range of energies is not known in advance, the 2D Ising model presents a unique challenge since the energy range is readily known and easily discovered by WL methods.
\end{abstract}

@@

\section{Introduction}
Flat-histogram Monte-Carlo simulation algorithms calculate the thermodynamic
properties of various systems over a range of temperatures.  The first histogram
method used a single canonical Monte Carlo simulation to predict properties for
nearby temperatures~\cite{ferrenberg1988new}. While the method effectively
samples a narrow energy range given limited by a narrow temperature range, it
proves computationally inefficient at sampling large energy ranges.

Multicanonical methods, introduced by Berg and Neuhaus, enabled flat-histogram
sampling which improved the exploration of configuration space and allowed the
system to overcome free-energy barriers~\cite{berg1991multicanonical, berg1992multicanonical}.
These works led to increase in the development of a variety of ``flat'' (or ``broad'') histogram
methods~\cite{penna1996broad, penna1998broad, swendsen1999transition,
wang2001determining, wang2001efficient} which could explore a wider range
of energies.  In addition to obtaining thermodynamic information for the entire
energy range for a single simulation, these approaches cannot be easily trapped
in a local energy minimum like a canonical simulation.

Wang and Landau introduced one of the most widely used flat-histogram
Monte-Carlo algorithms that accurately determined the density of states (DOS) for
a statistical system~\cite{wang2001determining,wang2001efficient}. For all of
its power, the method unfortunately requires \emph{a priori.} knowledge of several
user-defined parameters. Thus, for any given system under study, the user needs
to determine the ideal parameters in order to apply the method. The Wang-Landau
algorithm is also known to violate detailed balance (although only for brief
time intervals)~\cite{yan2003fast, shell2002generalization}. With the violation
of detailed balance, convergence of the algorithm is not guaranteed.

Because of the uncertainty of convergence for WL, many studies were undertaken
to understand how the modification (or update) factor $\gamma$ impacts the
convergence~\cite{zhou2005understanding,lee2006convergence,
belardinelli2007wang}. Belardinelli and Pereyra showed that an update factor that decreases faster than $1/t$ leads to nonconvergence~\cite{belardinelli2007wang, belardinelli2008analysis, zhou2008optimal}, where $t$ corresponds to the number of moves. Schneider \emph{et al.} outline minor refinements algorithm including scaling the update factor with the number of energy bins~\cite{schneider2017convergence}.
These studies led to the
formation of the $1/t$-WL algorithm and also sparked some debate as to whether
the entire simulation should be split into multiple parts, called ``production run" WL, in order to preserve ergodicity and
detailed balance~\cite{jayasri2005wang, mukhopadhyay2008monte}.

Liang independently considered whether WL could be treated as a special case of
Stochastic Approximation whose convergence could be mathematically
proven~\cite{liang2006theory, liang2007stochastic}. In 2007, Liang et
al.~\cite{liang2007stochastic} argued that WL can be considered a form of
Stochastic Approximation Monte Carlo (SAMC). Unlike WL, SAMC can guarantee
convergence (if certain conditions are met). Despite the added benefit of
guaranteed convergence, the method still has a system specific user-defined
variable. Such variables often create difficulty when applying Monte-Carlo
methods across arbitrary systems.

@@

The SAD (Stochastic Approximation with
Dynamic $\gamma$) method outlined by Pommerenck
et.al~\cite{pommerenck2020stochastic} is a special version of the SAMC algorithm
that dynamically chooses the modification factor rather than relying on system
dependent parameters. SAD shares the same convergence properties with SAMC while
replacing un-physical user-defined parameters with the algorithms dynamic
choice.

@@

\section{Ising Model}
The 2D Ising spin-lattice system is widely used as a testbed when
benchmarking or comparing Monte-Carlo
methods~\cite{ferdinand1969bounded, wang1999transition, trebst2004optimizing, barash2019estimating}. The 2nd order
phase transition behavior and the ability to directly calculate the
exact solution for finite lattices~\cite{beale1996exact, haggkvist2004computation} make the
system sufficiently interesting for such theoretical comparisons. It is
also important to note that direct comparison of the other methods can
be made with WL as its original implementation was done on this
system~\cite{wang2001determining,wang2001efficient}.
We test the convergence of several flat-histogram methods
on the periodic 2D square lattice ferromagnetic Ising model with nearest
neighbor interactions~\cite{landau2004new}.
\begin{align}
\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j - h \sum_i s_i
\end{align}
The $N\times N$ spin system can take on values of $\sigma_i = \pm 1$
for up or down spins respectively. In the absence of a magnetic field ($h =
0$), We can write the Hamiltonian as follows~\cite{onsager1944crystal,
kaufman1949crystal}:
\begin{align}
\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j
\end{align}
where the sum is over nearest neighbor spin sites. Beale showed that for finite
lattices the Density of States could directly be calculated from the partition
function~\cite{beale1996exact}
\begin{align}
Z = \sum_E g(E) e^{-{\beta E}}
\end{align}
% where $g(E)$ is the multiplicity of the system which is proportional to
% $D(E)$. We can estimate the average deviation from the exact solution
% using the relation~\cite{schneider2017convergence, shakirov2018convergence,
% barash2017control}:
% \begin{align}
% \braket{\epsilon(t)}_E = \frac{1}{N_E - 1} \sum_E \left( \frac{S(E,t) - S_{\text{Beale}(E)}}{S_{\text{Beale}(E)}} \right)
% \end{align}
where $g(E)$ is the multiplicity of the system which is proportional to
$D(E)$. We can estimate the maximum deviation in the canonical specific heat capacity $c_V$ from the exact solution~\cite{schneider2017convergence, shakirov2018convergence,
barash2017control,barash2017gpu}:
\begin{align}
\braket{\epsilon(t)}_E = \left|\max_E \left(c_V(E,S,T,t) - c_V(E,S,T,t)_{\text{Beale}}\right)\right|
\end{align}
% As per the implementation by Wang and
% Landau~\cite{wang2001determining}, the total number of available energy
% states are $N-1$.
Computing the specific heat capacity $c_V$ presents a difficult challenge for
any Monte-Carlo method due to fluctuations in the derivative of the internal energy.  Methods that accurately compute $c_V$ also by extension accurately compute the internal energy.

@@

Flat-histogram methods employ a random walk in energy space to estimate $D(E)$.  Each method operates by continuously updating the weights at each
step of the simulation
\begin{equation}
	\ln{w_{t+1}(E)}=\ln{w_{t}(E)}
	+\gamma_t
\end{equation}
where $t$ is number of the current move, $\gamma(t)$ is a move-dependent update
factor, and $E$ is the current energy.  This update causes the random walk to
avoid frequently sampling the same energies, leading to a rapid exploration
of energy space. Flat-histogram methods differ primarily in how they schedule
the decrease of $\gamma_t$.

@@

The WL method can alternatively be implemented as part of a production run which aids in
equilibration~\cite{gross2018massively}. WL is used to generate the weights
resulting in a flat-histogram~\cite{janke2017generalized}. Once a user-defined
minimum $\gamma^{\text{WL}}_t$ is reached, the simulation switches to a final production run
with fixed weights. WL implemented in this way can ensure ergodicity and
detailed-balance; however, the convergence of simulation is still impacted by
the choice of the minimum $\gamma^{\text{WL}}_t$.

Another weight-based flat-histogram method is
the Stochastic Approximation Monte Carlo (SAMC) algorithm. SAMC has a simple
schedule by which the update factor $\gamma^{\text{SA}}_t$ is continuously
decreased~\cite{liang2007stochastic, werlich2015stochastic,
schneider2017convergence}.  The update factor is defined in the
original implementation~\cite{liang2007stochastic} in terms of an
arbitrary tunable parameter $t_0$,
\begin{align}
\gamma_{t}^{\text{SA}} =\frac{t_0}{\max(t_0,t)}
\end{align}
where as above $t$ is the number of moves that have been attempted.

@@

Pommerenck et al. propose a refinement~\cite{pommerenck2020stochastic} to SAMC
where the update factor is determined dynamically rather than by the user.
Stochastic Approximation with a dynamic $\gamma$ (SAD) requires the user to
provide the lowest temperature of interest $T_{\min}$. This is analogous to WL
requiring \emph{a priori.} an energy range of interest; however, this is almost always easier to
identify and is more physical than the SAMC parameter $t_0$. The update factor
can be written in terms of the current estimates for the highest $E_H$ and
lowest $E_L$ energies of interest and the last time that an energy in the range
of interest is encountered $t_L$.
\begin{align}
  \gamma_{t}^{\text{SAD}} =
     \frac{
       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{t_L}
     }{
       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{N_S}\frac{t}{t_L}
     }
\end{align}
SAD only explores the energy range of interest as specified by the minimum
temperature of interest $T_{\min} < T < \infty$. During the simulation the two
energies $E_H$ and $E_L$, are refined such that the range of energies are conservatively
estimated. The weights are calculated for each energy region according to the original
prescription.
\begin{enumerate}
\item {$E < E_L$:} $w(E>E_H) = w(E_H)$
\item {$E_L < E < E_H$:} moves are handled the same as other weight-based
methods that are mentioned
\item {$E > E_H$:} $w(E<E_L) = w(E_L)e^{-\frac{E_L-E}{T_{\min}}}$
\end{enumerate}
Each time the simulation changes the value of $E_H$ or $E_L$, the weights
within the new portion of the interesting energy range are updated.

\section{Results}
We tested the algorithms on two different system sizes of the 2D Ising model.  The first is a smaller
simulation with a lattice size of $N = 32 \times 32$ and the second has a lattice size of $N = 128 \times 128$. The SAD method explores the energy
space of the each system using a minimum reduced temperature of $T_{\text{min}} = 1$. All simulations lead to the minimum important energy $E_{\min}$
and maximum entropy energy $E_{\max}$ being calculated (with the exception of
the WL methods where both of these parameters are needed \emph{a priori.}).

@@

\subsection{The 32 $\times$ 32 Ising model}
Fig.~\ref{fig:N32-gamma} shows the update factor $\gamma_t$ for each of the flat-histogram methods. All of the update factors initially start at $\gamma_{t=0} = 1$. SAD dynamically updates $\gamma_t$ throughout the simulation. After about $10^{10}$ moves, $\gamma^{\text{SAD}}_t$ proceeds as $1/t$. Both WL and ``production'' WL update factors $\gamma^{\text{WL}}_t$ are shown. The production WL begins after $\gamma^{\text{WL}}_t$ has reached $10^{-4}$. The update factor for $1/t$-WL decreases similarly to WL before finding all the energy states and switching to $1/t$. All of the SAMC update factors equal 1 until the number of moves is equal to $t_0$ at which point the decrease as $1/t$.

@@

\subsection{128 $\times$ 128 Ising system}
Fig.~\ref{fig:N128-gamma} shows the update factor $\gamma_t$ for each of the flat-histogram methods. All of the update factors initially start at $\gamma_{t=0} = 1$. SAD dynamically updates $\gamma_t$ throughout the simulation. After about $10^{13}$ moves, $\gamma^{\text{SAD}}_t$ proceeds as $1/t$. Both WL and ``production'' WL update factors $\gamma^{\text{WL}}_t$ are shown. The production WL begins after $\gamma^{\text{WL}}_t$ has reached $10^{-4}$ and $10^{-6}$. The update factor for $1/t$-WL decreases similarly to WL before finding all the energy states and switching to $1/t$. All of the SAMC update factors equal 1 until the number of moves is equal to $t_0$ at which point the decrease as $1/t$.

Fig.~\ref{fig:N128-cv-error} shows the maximum error in the heat capacity as a function of time for
this system. The solid/dashed lines
represent the average of the maximum value of the error in the specific heat $c_V$ averaged
over eight simulations using different random number seeds. The range of maximum
errors for each simulation is shown as a shaded region. By
the time $10^13$ moves have been made all but the WL simulation have begun to
converge as $1/\sqrt{t}$. We then see the WL error saturate around $10^12$ moves.

@@

We find that both SAD, 1/t-WL, and production WL (provided an ideal minimum update factor before beginning the production run) demonstrate excellent and robust convergence.
They all converge more rapidly than SAMC, and unlike WL do not suffer from
error saturation. We find that for larger systems SAD reduces the updated factor
more slowly (and conservatively) than 1/t-WL and production WL. This means that SAD will take
proportionately more moves to converge to the same value as 1/t-WL as system
size is increased. SAD requires the user to specify a temperature range of
interest rather than an energy range of interest as the WL methods do. The 2D Ising model is a system where the energy range is known \emph{a priori.}, which makes a comparison between WL methods and SAD (which requires a minimum temperature) more biased towards WL methods. For the majority of cases in which the energy range of the system is not known and where a range of desired temperatures is known, this will make the SAD method
considerably more convenient. This work also visually demonstrates that production WL performs extremely well and is preferable to WL for ensuring both ergodicity and detailed-balance.

@@

We thank Klas Markstr\"{o}m for discussions regarding the exact computation of the partition function for the 2D Ising model. We also with to thank Johannes Zierenberg for helpful discussions regarding WL production run simulations and
Martin Weigel for discussions regarding entropic population annealing.

>>>>

\begin{abstract}
We examine the convergence of the stochastic approximation with a dynamic
update factor (SAD) algorithm applied to the 2D Ising model. Comparison with
stochastic approximation Monte Carlo (SAMC) and Wang-Landau (WL) methods show that SAD performs robustly and without user input knowledge of an energy range. We also compare SAD with ``production'' run WL with fixed weights which is the first of its kind recorded. Since SAD is more powerful in the common case in which the range of energies is not known in advance, the 2D Ising model presents a unique challenge since the energy range is readily known and easily discovered by WL methods.
\end{abstract}

@@

\section{Introduction}
Flat-histogram Monte Carlo simulation algorithms calculate the thermodynamic
properties of various systems over a range of temperatures.  The first histogram
method used a single canonical Monte Carlo simulation to predict properties for
nearby temperatures~\cite{ferrenberg1988new}. While the method effectively
samples a narrow energy range, it proves computationally inefficient at sampling large energy ranges.

Multicanonical methods, introduced by Berg and Neuhaus, enabled flat-histogram
sampling which improved the exploration of configurational space and allowed the
simulation to overcome free-energy barriers~\cite{berg1991multicanonical, berg1992multicanonical}.
These works led to increase in the development of a variety of ``flat'' (or ``broad'') histogram
methods~\cite{penna1996broad, penna1998broad, swendsen1999transition,
wang2001determining, wang2001efficient} which could explore a wider range
of energies.  In addition to obtaining thermodynamic information for the entire
energy range for a single simulation, these approaches cannot be easily trapped
in a local energy minimum like a canonical simulation.

Wang and Landau introduced one of the most widely used flat-histogram
Monte Carlo algorithms that determined the density of states (DOS) for
a statistical system~\cite{wang2001determining,wang2001efficient}. For all of
its power, the method unfortunately requires \emph{a priori.} knowledge of several
user-defined parameters. Thus, for any given system under study, the user needs
to determine the ideal parameters in order to apply the method. The Wang-Landau
algorithm is also known to violate detailed balance (although only for brief
time intervals)~\cite{yan2003fast, shell2002generalization}. With the violation
of detailed balance, convergence of the algorithm is not guaranteed.

Because of the uncertainty of convergence for WL, many studies have been undertaken
to understand how the modification (or update) factor $\gamma$ impacts the
convergence~\cite{zhou2005understanding,lee2006convergence,
belardinelli2007wang}. Belardinelli and Pereyra showed that an update factor that decreases faster than $1/t$ leads to nonconvergence~\cite{belardinelli2007wang, belardinelli2008analysis, zhou2008optimal}, where $t$ corresponds to the number of moves. Schneider \emph{et al.} outline minor refinements algorithm including scaling the update factor with the number of energy bins~\cite{schneider2017convergence}.
These studies led to the
formation of the $1/t$-WL algorithm and also sparked some debate as to whether
the entire simulation should be split into multiple parts, called ``production run" WL, in order to preserve ergodicity and
detailed balance~\cite{jayasri2005wang, mukhopadhyay2008monte}.

Liang independently considered whether WL could be treated as a special case of
stochastic approximation whose convergence could be mathematically
proven~\cite{liang2006theory, liang2007stochastic}. In 2007, Liang et
al.~\cite{liang2007stochastic} argued that WL can be considered a form of
stochastic approximation Monte Carlo (SAMC). Unlike WL, SAMC can guarantee
convergence (if certain conditions are met). Despite the added benefit of
guaranteed convergence, the method still has a system specific user-defined
variable. Such variables often create difficulty when applying Monte Carlo
methods across arbitrary systems.

@@

The SAD (stochastic approximation with a
dynamic $\gamma$) method outlined by Pommerenck
et.al~\cite{pommerenck2020stochastic} is a special version of the SAMC algorithm
that dynamically chooses the modification factor rather than relying on system
dependent parameters. SAD shares the same convergence properties with SAMC while
replacing un-physical user-defined parameters with the algorithms dynamic
choice.

@@

\section{Ising Model}
The 2D Ising spin-lattice system is widely used as a testbed when
benchmarking or comparing Monte Carlo
methods~\cite{ferdinand1969bounded, wang1999transition, trebst2004optimizing, barash2019estimating}. The 2nd order
phase transition behavior and the ability to directly calculate the
exact solution for finite lattices~\cite{beale1996exact, haggkvist2004computation} make the
system sufficiently interesting for such theoretical comparisons. It is
also important to note that direct comparison of the other methods can
be made with WL as its original implementation was done on this
system~\cite{wang2001determining,wang2001efficient}.
We test the convergence of several flat-histogram methods
on the periodic 2D square lattice ferromagnetic Ising model with nearest
neighbor interactions~\cite{landau2004new}.
\begin{align}
\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j - h \sum_i s_i
\end{align}
The $N\times N$ spin system can take on values of $\sigma_i = \pm 1$
for up or down spins respectively. In the absence of a magnetic field ($h =
0$), We can write the Hamiltonian as follows~\cite{onsager1944crystal,
kaufman1949crystal}:
\begin{align}
\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j
\end{align}
where the sum is over nearest neighbor spin sites. Beale showed that for finite
lattices the Density of States could directly be calculated from the partition
function~\cite{beale1996exact}
\begin{align}
Z = \sum_E g(E) e^{-{\beta E}}
\end{align}
% where $g(E)$ is the multiplicity of the system which is proportional to
% $D(E)$. We can estimate the average deviation from the exact solution
% using the relation~\cite{schneider2017convergence, shakirov2018convergence,
% barash2017control}:
% \begin{align}
% \braket{\epsilon(t)}_E = \frac{1}{N_E - 1} \sum_E \left( \frac{S(E,t) - S_{\text{Beale}(E)}}{S_{\text{Beale}(E)}} \right)
% \end{align}
where $g(E)$ is the multiplicity of the system which is proportional to
$D(E)$. We can estimate the maximum deviation in the canonical specific heat capacity $c_V$ from the exact solution~\cite{schneider2017convergence, shakirov2018convergence,
barash2017control,barash2017gpu}:
\begin{align}
\braket{\epsilon(t)}_E = \left|\max_E \left(c_V(E,S,T,t) - c_V(E,S,T,t)_{\text{Beale}}\right)\right|
\end{align}
% As per the implementation by Wang and
% Landau~\cite{wang2001determining}, the total number of available energy
% states are $N-1$.
Computing the specific heat capacity $c_V$ presents a difficult challenge for
any Monte Carlo method due to fluctuations in the derivative of the internal energy.  Methods that accurately compute $c_V$ also by extension accurately compute the internal energy.

@@

Flat-histogram methods employ a random walk in energy space to estimate $D(E)$.  Each method operates by continuously updating the weights at each
step of the simulation
\begin{equation}
	\ln{w_{t+1}(E)}=\ln{w_{t}(E)}
	+\gamma_t
\end{equation}
where $t$ is number of the current move, $\gamma(t)$ is a move-dependent update
factor, and $E$ is the current energy.  This update causes the random walk to
avoid frequent sampling of the same energies, leading to a rapid exploration
of energy space. Flat-histogram methods differ primarily in how they schedule
the decrease of $\gamma_t$.

@@

The WL method can alternatively be implemented as part of a production run which aids in
equilibration~\cite{gross2018massively}. WL is used to generate the weights
resulting in a flat-histogram~\cite{janke2017generalized}. Once a user-defined
minimum $\gamma^{\text{WL}}_t$ is reached, the simulation switches to a final production run
with fixed weights. WL implemented in this way can ensure ergodicity and
detailed balance; however, the convergence of simulation is still impacted by
the choice of the minimum $\gamma^{\text{WL}}_t$.

Another weight-based flat-histogram method is
the stochastic approximation Monte Carlo (SAMC) algorithm. SAMC has a simple
schedule by which the update factor $\gamma^{\text{SA}}_t$ is continuously
decreased~\cite{liang2007stochastic, werlich2015stochastic,
schneider2017convergence}.  The update factor is defined in the
original implementation~\cite{liang2007stochastic} in terms of an
arbitrary tunable parameter $t_0$,
\begin{align}
\gamma_{t}^{\text{SA}} =\frac{t_0}{\max(t_0,t)}
\end{align}
where as above $t$ is the number of moves that have been attempted.

@@

Pommerenck et al. propose a refinement~\cite{pommerenck2020stochastic} to SAMC
where the update factor is determined dynamically rather than by the user.
stochastic approximation with a dynamic $\gamma$ (SAD) requires the user to
provide the lowest temperature of interest $T_{\min}$. This is analogous to WL
requiring \emph{a priori.} an energy range of interest; however, this is almost always easier to
identify and is more physical than the SAMC parameter $t_0$. The update factor
can be written in terms of the current estimates for the highest $E_H$ and
lowest $E_L$ energies of interest and the last time that an energy in the range
of interest is encountered $t_L$.
\begin{align}
  \gamma_{t}^{\text{SAD}} =
     \frac{
       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{t_L}
     }{
       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{N_S}\frac{t}{t_L}
     }
\end{align}
SAD only explores the energy range of interest as specified by the minimum
temperature of interest $T_{\min} < T < \infty$. During the simulation the two
energies $E_H$ and $E_L$, are refined such that the range of energies are conservatively
estimated. The weights are calculated for each energy region according to the original
prescription.
\begin{enumerate}
\item {$E < E_L$:} $w(E>E_H) = w(E_H)$
\item {$E_L < E < E_H$:} moves are handled the same as other weight-based
methods that are mentioned
\item {$E > E_H$:} $w(E<E_L) = w(E_L)e^{-\frac{E_L-E}{T_{\min}}}$
\end{enumerate}
Each time the simulation changes the value of $E_H$ or $E_L$, the weights
within the new portion of the interesting energy range are updated.

\section{Results}
We tested the algorithms on two different system sizes of the 2D Ising model.  The first is a smaller
simulation with a lattice size of $N = 32 \times 32$ and the second has a lattice size of $N = 128 \times 128$. The SAD method explores the energy
space of each system using a minimum reduced temperature of $T_{\text{min}} = 1$. All simulations lead to the minimum important energy $E_{\min}$
and maximum entropy energy $E_{\max}$ being calculated (with the exception of
the WL methods where both of these parameters are needed \emph{a priori.}).

@@

\subsection{The 32 $\times$ 32 Ising model}
Fig.~\ref{fig:N32-gamma} shows the update factor $\gamma_t$ for each of the flat-histogram methods. All of the update factors initially start at $\gamma_{t=0} = 1$. SAD dynamically updates $\gamma_t$ throughout the simulation. After about $10^{10}$ moves, $\gamma^{\text{SAD}}_t$ proceeds as $1/t$. Both WL and ``production'' WL update factors $\gamma^{\text{WL}}_t$ are shown. The production WL begins after $\gamma^{\text{WL}}_t$ has reached $10^{-4}$. The update factor for $1/t$-WL decreases similarly to WL before finding all the energy states and switching to $1/t$. All of the SAMC update factors equal 1 until the number of moves is equal to $t_0$ at which point they decrease as $1/t$.

@@

\subsection{128 $\times$ 128 Ising system}
Fig.~\ref{fig:N128-gamma} shows the update factor $\gamma_t$ for each of the flat-histogram methods. All of the update factors initially start at $\gamma_{t=0} = 1$. SAD dynamically updates $\gamma_t$ throughout the simulation. After about $10^{13}$ moves, $\gamma^{\text{SAD}}_t$ proceeds as $1/t$. Both WL and ``production'' WL update factors $\gamma^{\text{WL}}_t$ are shown. The production WL begins after $\gamma^{\text{WL}}_t$ has reached $10^{-4}$ and $10^{-6}$. The update factor for $1/t$-WL decreases similarly to WL before finding all the energy states and switching to $1/t$. All of the SAMC update factors equal 1 until the number of moves is equal to $t_0$ at which point they decrease as $1/t$.

Fig.~\ref{fig:N128-cv-error} shows the maximum error in the heat capacity as a function of time for
this system. The solid/dashed lines
represent the average of the maximum value of the error in the specific heat $c_V$ averaged
over eight simulations using different random number seeds. The range of maximum
errors for each simulation is shown as a shaded region. By
the time $10^{13}$ moves have been made all but the WL simulation have begun to
converge as $1/\sqrt{t}$. We then see the WL error saturate around $10^{12}$ moves.

@@

We find that both SAD, 1/t-WL, and production WL (provided an ideal minimum update factor before beginning the production run) demonstrate excellent and robust convergence.
They all converge more rapidly than SAMC, and unlike WL do not suffer from
error saturation. We find that for larger systems SAD reduces the updated factor
more slowly (and conservatively) than 1/t-WL and production WL. This means that SAD will take
proportionately more moves to converge to the same value as 1/t-WL as system
size is increased. SAD requires the user to specify a temperature range of
interest rather than an energy range of interest as the WL methods do. The 2D Ising model is a system where the energy range is known \emph{a priori.}, which makes a comparison between WL methods and SAD (which requires a minimum temperature) more biased towards WL methods. For the majority of cases in which the energy range of the system is not known and where a range of desired temperatures is known, this will make the SAD method
considerably more convenient. This work also visually demonstrates that ``production'' WL performs extremely well and is preferable to WL for ensuring both ergodicity and detailed balance.

@@

We thank Klas Markstr\"{o}m for discussions regarding the exact computation of the partition function for the 2D Ising model. We also wish to thank Johannes Zierenberg for helpful discussions regarding WL production run simulations and
Martin Weigel for discussions regarding entropic population annealing.
